A fault tolerant multi-server system for MMOGs

### 2011-07-07 trunk/textos/phd-prospectus/prospectus.tex
prospectus ready for being reviewed by the advisor

\abstract {
MMOGs -- massively multiplayer online games -- have become, in the last decade, an important genre of online entertainment, having a significant market share. In these games, many thousands of participants play simultaneously with one another in the same match. However, this kind of game is traditionally supported by a powerful (and expensive) central infrastructure with considerable cpu power and very fast and low latency Internet connections. This work proposes the use of a geographically distributed server system, formed by voluntary nodes which help serving the game to the players. Two very important features, therefore, must be provided: consistency (which may be week and only existing in the players' perspective) and fault tolerance, as some of the participating nodes may crash and become unavailable. For this work, we consider the crash-stop failure model.
}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Optimistic multicast} \label{sec:optmcast}

As the servers are organized in groups, and every command is sent to one or more of such groups, we can use a multicast primitive. As commands seen by different servers must be handled in the same order, we need total order. As the commands sent by each player should be executed in the same order they were issued in, the multicast should provide FIFO ordering.

However, we want to minimize the inter-group communication overhead, so that our distributed server would be more scalable. Non-genuine multicast primitives \cite{schiper2009gng} require that control messages keep being exchanged between groups, even when there is no application level message to be sent. On the other hand, although genuine multicast algorithms minimize the number of messages exchanged, they impose a greater latecy on their delivery \cite{schiper2008ica}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future work} \label{sec:futurework}
%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Activities and publications} \label{sec:achievements}
%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\abstract {
MMOGs -- massively multiplayer online games -- have become, in the last decade, an important genre of online entertainment, having a significant market share. In these games, many thousands of participants play simultaneously with one another in the same match. However, this kind of application is traditionally supported by a powerful (and expensive) central infrastructure with considerable cpu power and very fast and low latency Internet connections. This work proposes the use of a geographically distributed server system, formed by nodes which help serving the game to the players. The most important features that must be provided are: consistency (since different players should have the sense of one single shared virtual environment), timeliness (as these games are usually real-time applications) and fault tolerance (as some of the participating nodes may crash and become unavailable).
}

@@

% Figure \ref{fig:consistency} illustrates two avatars situated in regions managed by different server groups, but still close enough to interact. Each command issued by their players has to be forwarded to the groups that manage some object in its target set, which is done by multicast.
% 
% \begin{figure}[!h]
%   \centering
%   \includegraphics[width=0.8\linewidth]{images/consistency}
%   \caption{Multicasting a command to a group which manages an object in that command's target set}
%   \label{fig:consistency}
% \end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Quasi-genuine multicast} \label{sec:optmcast}

As the servers are organized in groups, and every command is sent to one or more of such groups, we can use a multicast protocol. Also, as commands seen by different servers must be handled in the same order, we need total order, so that the state machine replication approach can be used. Besides, the commands sent by each player should be executed in the same order in which this player issued them; therefore, the multicast protocol should provide FIFO ordering.

However, the inter-group communication overhead should be minimized, so that the distributed server would be more scalable. Non-genuine multicast primitives \cite{schiper2009gng} require that control messages keep being exchanged between groups, even when there is no application level message to be sent. On the other hand, although genuine multicast algorithms minimize the number of messages exchanged, they impose a greater latecy on their delivery \cite{schiper2008ica}.

We devised, then, a multicast protocol that, although not genuine, allows the system to increase its scale without getting flooded with control messages. We call it \emph{quasi-genuine}, because every group keeps sending messages to (and receiving from) other groups, even if there is no application level message to be exchanged, but only the ones with which it \emph{is possible to} exchange messages. The information concerning which other groups a given group can communicate with is given by the application.

The advantage of a quasi-genuine multicast protocol is that it would be a middle ground between genuine and non-genuine protocols, inheriting some advantages from both -- namely, improved scalability and lower message delivery time --, at the cost of restricting which groups can communicate with each other. Although the system can be reconfigured to change the group pairs that are able to exchange messages, such reconfiguration will have some cost, so this multicast protocol is more suitable for applications where this would not be frequent.

This can be used by our consistency management algorithm because every server group only has to forward commands to a specific set of other groups, which are the ones managing neighbor regions, and this is known as soon as the partitioning has been performed. Changes to the partioning should be performed rarely, as it would incur in some players migrating between servers. Therefore, the set of group pairs that can communicate with each other will not change frequently.

To further decrease the message delivery time, an optimistic delivery will also be employed. By assigning timestamps based on the local clocks -- which are supposed to be more or less synchronized -- and by every server waiting long enough for messages to arrive from another servers before performing the delivery, the final total message order will most likely match the timestamp order, so that the messages could be delivered in only one communication step, although that delivery would have to be confirmed by some conservative delivery.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Published work} \label{sec:achievements}

Some related work that we have published so far include:
\begin{itemize}
  \item C. E. B. Bezerra, J. L. D. Comba, C. F. R. Geyer. A fine granularity load balancing technique for MMOG servers using a kd-tree to partition the space. \textit{VIII Brazilian Symposium on Computer Games and Digital Entertainment - Computing Track}, 2009.
  \item C. E. B. Bezerra, C. F. R. Geyer. A load balancing scheme for massively multiplayer online games. \textit{Multimedia Tools and Applications - Special Issue on Massively Multiplayer Online Gaming Systems and Applications}, v. 1, n. 1, pp. 263-289, 2009.
  \item C. E. B. Bezerra, F. R. Cecin, C. F. R. Geyer. A3: a novel interest management algorithm for distributed simulations of MMOGs. \textit{12-th IEEE International Symposium on Distributed Simulation and Real Time Applications}, 2008.
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future work} \label{sec:futurework}

We are currently workings towards the validation of both the quasi-genuine multicast and the game state consistency management protocols, which require both formal proofs regarding correctness and experimental results to demonstrate their performance. Another direction that we intend to follow is to consider not only crash-stop faults, but also Byzantine faults. By tolerating these faults, the server system would also be resistant to cheating, which can also be seen as Byzantine behavior.

%%%%%%%%%%%%%%%%%%%%%%%%%
